### 2022-07-19 paper/gradVsStokesSolvesPlot.tex
standalone tikz files to be used with \include to generate various tables and tikz figures

\begin{figure}[H]
\begin{center} 
\begin{tikzpicture}[baseline, trim axis right]
\begin{axis}[width=0.99\textwidth, height=0.75\textwidth,
            ymode=log,
            ymin=5e-9,
            ymax=1,
            xmax=900,
            xtick={0, 100, 200, 300, 400, 500, 600, 700, 800, 900},
            xlabel={Cumulative Stokes solves},
            ylabel={$||g_{k}||_{2}/||g_{0}||_{2}$, relative norm of the gradient},
            title={Convergence of the preconditioned inexact Newton-CG},
			legend pos=north east,
]
\addlegendentry{None}
\addplot[dash dot, line width=0.5mm] table {gradVsCumulativeStokes_none.dat};
\addlegendentry{Reg}
\addplot[dotted, line width=0.5mm] table {gradVsCumulativeStokes_reg.dat};
\addlegendentry{PSF (5)}
\addplot[red, line width=0.5mm] table {gradVsCumulativeStokes_psf.dat};
\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}