### 2022-07-20 paper/PreconditionedSpectrumPlot.tex

\begin{figure}[H]
\begin{center} 
\begin{tikzpicture}[baseline, trim axis right]
\begin{axis}[width=0.99\textwidth, height=0.75\textwidth,
            ymode=log,
            xmin=-50,
            xmax=1450,
            xlabel={$k$, generalized eigenvalue \#},
            ylabel={$\lambda_{k}$},
            xtick={0, 200, 400, 600, 800, 1000, 1200, 1400},
			legend pos=north east,
			title={Generalized eigenvalues, $Hu_{k}=\lambda_{k}Bu_{k}$}
]
\addlegendentry{REG}
\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
\addlegendentry{PSF (1)}
\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
\addlegendentry{PSF (5)}
\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
\addlegendentry{PSF (25)}
\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}