### 2022-07-20 paper/RegularizationRobustnessPlot.tex

\begin{figure}[H]
\begin{center} 
\begin{tikzpicture}[baseline, trim axis right]
\begin{axis}[width=0.99\textwidth, height=0.75\textwidth,
            xmode=log,
            ymode=log,
            xlabel={$\gamma$, regularization parameter},
            ylabel={\# of CG iterations},
            %legend pos=north east,
			title={CG iterations to achieve $||H x-b||/||b||<10^{-6}$}
]
\addlegendentry{REG}
\addplot[dotted, line width=0.5mm] table[x=gamma, y=reg] {CGitsvsGamma.dat};
\addlegendentry{None}
\addplot[dash dot, line width=0.5mm] table[x=gamma, y=none] {CGitsvsGamma.dat};
\addlegendentry{PSF (1)}
\addplot[blue, line width=0.5mm] table[x=gamma, y=psf1] {CGitsvsGamma.dat};
\addlegendentry{PSF (5)}
\addplot[red, line width=0.5mm] table[x=gamma, y=psf5] {CGitsvsGamma.dat};
\addlegendentry{PSF (25)}
\addplot[purple, line width=0.5mm] table[x=gamma, y=psf25] {CGitsvsGamma.dat};
\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}