### 2023-06-01 paper/stokes_pcg_convergence.tex

\documentclass{standalone}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{amsmath}

% SIAM Shared Information Template
% This is information that is shared between the main document and any
% supplement. If no supplement is required, then this information can
% be included directly in the main document.

% Packages and macros go here
\usepackage{lipsum}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{bm}

\begin{document}
			\begin{tikzpicture}[baseline]%, trim axis right]
			\begin{axis}[width=0.49\linewidth, height=0.5\linewidth,
	ymode=log,
	ymin=1e-13,
	ymax=2,
	xmax=750,
	axis x line*=bottom,
	axis y line*=left,
	xtick={0, 200, 400, 600},
	title={CG convergence},
	xlabel={$k$, CG iteration},
	ylabel={$\|\bm{x}_k-\bm{x}\|/\|\bm{x}\|$},
	legend style={at={(1.2,1)},anchor=north east,draw=none},
	%legend pos=north east,
	legend style={font=\small}
	]
	\addlegendentry{Reg}		
	\addplot[dotted, line width=0.5mm] table {cg_relres_reg_0.05noise.dat};
	\addlegendentry{None}
	\addplot[dash dot, line width=0.5mm] table {cg_relres_none_0.05noise.dat};
	\addlegendentry{PSF (1)}
	\addplot[blue, line width=0.5mm] table {cg_relres_psf1_0.05noise.dat};
	\addlegendentry{PSF (5)}
	\addplot[red, line width=0.5mm] table {cg_relres_psf5_0.05noise.dat};
	\addlegendentry{PSF (25)}
	\addplot[purple, line width=0.5mm] table {cg_relres_psf25_0.05noise.dat};
	\end{axis}
	\end{tikzpicture}
\end{document}