### 2023-06-02 paper/stokes_pcg_convergence.tex
added new data for Stokes pcg convergence and generalized eigenvalues... updated the mini tex files which generate the associated tikz figures... this is for data in e6 Stokes data directory which indicates that the inverse problem was solved to a looser tolerance of requiring a 6 orders of magnitude in the reduction of the norm of the gradient

\documentclass{standalone}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{amsmath}

% SIAM Shared Information Template
% This is information that is shared between the main document and any
% supplement. If no supplement is required, then this information can
% be included directly in the main document.

% Packages and macros go here
\usepackage{lipsum}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{bm}

\begin{document}
			\begin{tikzpicture}[baseline]%, trim axis right]
			\begin{axis}[width=0.49\linewidth, height=0.5\linewidth,
	ymode=log,
	ymin=1e-13,
	ymax=2,
	xmax=750,
	axis x line*=bottom,
	axis y line*=left,
	xtick={0, 200, 400, 600},
	title={CG convergence},
	xlabel={$k$, CG iteration},
	ylabel={$\|\bm{x}_k-\bm{x}\|/\|\bm{x}\|$},
	legend style={at={(1.2,1)},anchor=north east,draw=none},
	%legend pos=north east,
	legend style={font=\small}
	]
	\addlegendentry{Reg}		
	\addplot[dotted, line width=0.5mm] table {Stokes_cg_rand_errs_reg_0.05noise.dat};
	\addlegendentry{None}
	\addplot[dash dot, line width=0.5mm] table {Stokes_cg_rand_errs_none_0.05noise.dat};
	\addlegendentry{PSF (1)}
	\addplot[blue, line width=0.5mm] table {Stokes_cg_rand_errs_psf1_0.05noise.dat};
	\addlegendentry{PSF (5)}
	\addplot[red, line width=0.5mm] table {Stokes_cg_rand_errs_psf5_0.05noise.dat};
	\addlegendentry{PSF (25)}
	\addplot[purple, line width=0.5mm] table {Stokes_cg_rand_errs_psf25_0.05noise.dat};
	\end{axis}
	\end{tikzpicture}
\end{document}