### 2023-06-05 paper/adv_pcg_convergence.tex
editing pass on paper with Tucker. Paper is ready for Noemi

\documentclass{standalone}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{amsmath}

% SIAM Shared Information Template
% This is information that is shared between the main document and any
% supplement. If no supplement is required, then this information can
% be included directly in the main document.

% Packages and macros go here
\usepackage{lipsum}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{bm}

\begin{document}
	\begin{tikzpicture}[baseline]%, trim axis right]
		\begin{axis}[width=0.49\linewidth, height=0.5\linewidth,
		ymode=log,
		ymin=1e-11,
		ymax=2,
		xmax=1000,
		axis x line*=bottom,
		axis y line*=left,
		xtick={0, 200, 400, 600,800,1000},
		xlabel={$j$, CG iteration},
		ylabel={$\|\bm{x}_j - \bm{x}\|/\|\bm{x}\|$},
		title={CG convergence},
		legend style={at={(1.2,1)},anchor=north east,draw=none},
		%legend pos=north east,
		legend style={font=\small}
		]
			\addlegendentry{Reg}
			\addplot[dotted, line width=0.5mm] table {advDiff_cg_rand_errs_reg.dat};
			\addlegendentry{None}
			\addplot[dash dot, line width=0.5mm] table {advDiff_cg_rand_errs_none.dat};
			\addlegendentry{PSF (1)}
			\addplot[blue, line width=0.5mm] table {advDiff_cg_rand_errs_psf1.dat};
			\addlegendentry{PSF (5)}
			\addplot[red, line width=0.5mm] table {advDiff_cg_rand_errs_psf5.dat};
			\addlegendentry{PSF (25)}
			\addplot[purple, line width=0.5mm] table {advDiff_cg_rand_errs_psf25.dat};
		\end{axis}
	\end{tikzpicture}
\end{document}