### 2023-07-06 paper_sisc_first_submission/KrylovConvergencePlot.tex
add subdirectory with frozen paper from first submission

\begin{figure}[H]
\begin{center} 
\begin{tikzpicture}[baseline, trim axis right]
\begin{axis}[width=0.99\textwidth, height=0.75\textwidth,
            ymode=log,
            ymin=1e-11,
            ymax=2,
            xmax=750,
            xtick={0, 100, 200, 300, 400, 500, 600, 700},
            xlabel={$k$, CG iteration},
            ylabel={$||r_{k}||_{2}/||r_{0}||_{2}$, relative residual norm},
			legend pos=north east,
]
\addlegendentry{None}
\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
\addlegendentry{Reg}
\addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
\addlegendentry{PSF (1)}
\addplot[blue, line width=0.5mm] table {cg_relres_p1.dat};
\addlegendentry{PSF (5)}
\addplot[red, line width=0.5mm] table {cg_relres_p5.dat};
\addlegendentry{PSF (25)}
\addplot[purple, line width=0.5mm] table {cg_relres_p25.dat};
\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}