### 2024-02-18 paper_sisc_first_revision/frog_tau_batches.tex
created tikz plots for frog convergence for different tau

\documentclass{standalone}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{amsmath}

% SIAM Shared Information Template
% This is information that is shared between the main document and any
% supplement. If no supplement is required, then this information can
% be included directly in the main document.

% Packages and macros go here
\usepackage{lipsum}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{bm}

\begin{document}
	\begin{tikzpicture}[baseline]%, trim axis right]
		\begin{axis}[width=0.49\linewidth, height=0.5\linewidth,
		xmode=log,
		ymode=log,
		ymin=1e-4,
		ymax=1e0,
%		xmax=1000,
		axis x line*=bottom,
		axis y line*=left,
%		xtick={0, 200, 400, 600,800,1000},
		xlabel={\#batches},
%		ylabel={$\|\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}\|_\text{Fro} / \|\boldsymbol{\Phi}\|_\text{Fro}$},
		title={PSF convergence (batches)},
%		legend style={at={(1.2,1)},anchor=north east,draw=none},
%		legend pos=south west,
%		legend style={font=\small}
		]
%			\addlegendentry{$\tau=2.0$}
			\addplot[black, line width=0.5mm] table {frog_convergence_batches_tau2.0.dat};
%			\addlegendentry{$\tau=2.5$}
			\addplot[blue, line width=0.5mm] table {frog_convergence_batches_tau2.5.dat};
%			\addlegendentry{$\tau=3.0$}
			\addplot[cyan, line width=0.5mm] table {frog_convergence_batches_tau3.0.dat};
%			\addlegendentry{$\tau=3.5$}
			\addplot[red, line width=0.5mm] table {frog_convergence_batches_tau3.5.dat};
%			\addlegendentry{$\tau=4.0$}
			\addplot[purple, line width=0.5mm] table {frog_convergence_batches_tau4.0.dat};
		\end{axis}
	\end{tikzpicture}
\end{document}