### 2024-02-19 revision/letter_revision.tex
added draft of response letter

\documentclass[11pt]{article}
\setlength{\hoffset}{-1in}       \setlength{\voffset}{-1in}
\setlength{\textwidth}{140mm}    \setlength{\textheight}{240mm}
\setlength{\topmargin}{18mm} \setlength{\oddsidemargin}{35mm}
\setlength{\evensidemargin}{25mm} \setlength{\headheight}{25mm}
\setlength{\headsep}{5mm} \setlength{\marginparwidth}{20mm}

\usepackage{graphicx}
\usepackage[right]{eurosym}

\begin{document}

\pagestyle{empty}
 \begin{figure}
 \begin{minipage}{.5\columnwidth}
 \end{minipage}\hfill
 \begin{minipage}[b]{.48\columnwidth}
\flushright
 \noindent{\sc\large Nick Alger}\vspace{-0.9cm}
 \end{minipage}
 \end{figure}
 \noindent\rule{\columnwidth}{.5pt}\vspace{-2ex}
 \begin{figure}[h]
 \begin{minipage}{.2\columnwidth}
 \end{minipage}\hfill
 \begin{minipage}{.78\columnwidth}
 \flushright\vspace{-2ex}
 {\noindent Oden Institute for Computational Engineering \& Sciences \\ 
 University of Texas at Austin, \\ 
 1 Univ.\ Station, C0200,
 TX 78712 USA \\
 email: {\tt nalger225@gmail.com}\\
 tel: +1 425 736 4111}\\
 \end{minipage}
 \end{figure}
% \vspace{1cm} \noindent \\
\phantom{adafdsf}
\vspace{3ex}

\bigskip

\noindent Dear Professor Antil:\\[2ex]

Please find enclosed the revised manuscript ``Point spread function
approximation of high rank Hessians with locally supported
non-negative integral kernels'' by Nick Alger, Tucker Hartland, Noemi
Petra, and Omar Ghattas which we would like to resubmit to SISC in the
Numerical Algorithms for Scientific Computing section.

%
This cover letter is accompanied by a point-by-point list of responses
to the referees' comments. In addition, we have prepared a diff
document that highlights the differences between the initially
submitted and the new manuscript, where revisions are indicated by
color (blue represents new text, and red is deleted text). Please send
the diff document to the reviewers as well.

We would like to thank the reviewers for their careful reading of our article 
and their constructive feedback, and many thanks to you for handling the
manuscript.\\[3ex]

\noindent Sincerely,\\[5ex]
\phantom{dd}\hspace{2cm} Nick Alger
\end{document}