### 2024-02-24 presentations/siamuq24-noemi/figures/stokes_geigs_figure.tex
noemis slides

\documentclass{standalone}
\usepackage{tikz}
\usepackage{pgfplots}

\begin{document}
	\begin{tikzpicture}[baseline]%, trim axis right]
		\begin{axis}[width=0.49\linewidth, height=0.5\linewidth,
		ymode=log,
		xmin=-50,
		xmax=1450,
		ymax=1.e7,
		title={Generalized eigenvalues},
		xlabel={$k$, generalized eigenvalue \#},
		ylabel={$\lambda_{k}$},
		xtick={0, 300, 600, 900, 1200},
		%legend pos=north east,
		legend style={font=\small},
		ylabel near ticks, ylabel shift={-6pt},
		axis x line*=bottom,
		axis y line*=right,
		%legend style={at={(0.95,1)},anchor=north east},
		]
		%\addlegendentry{REG}
		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {Stokes_generalizedEigenvalues_0.05noise.dat};
		\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {Stokes_generalizedEigenvalues_0.05noise.dat};
		%\addlegendentry{PSF (1)}
		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {Stokes_generalizedEigenvalues_0.05noise.dat};
		%\addlegendentry{PSF (5)}
		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {Stokes_generalizedEigenvalues_0.05noise.dat};
		%\addlegendentry{PSF (25)}
		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {Stokes_generalizedEigenvalues_0.05noise.dat};
		\end{axis}
	\end{tikzpicture}
\end{document}