### 2024-03-01 presentations/siampp24-tucker/localpsf_revised_figures/MorozovDiscrepancyPlot.tex

\begin{figure}[H]
\begin{center} 
\begin{tikzpicture}[baseline, trim axis right]
\begin{axis}[width=0.75\textwidth, height=0.75\textwidth,
            xmode=log,
            ymode=log,
            grid=both,
            xlabel={$\gamma$, regularization parameter},
			legend pos=north west,
]
\addlegendentry{data misfit}
\addplot[color=blue, mark=o, line width=0.5mm] table[x=morozov_gammas, y=morozov_discrepancies] {MorozovData.dat};
\addlegendentry{norm of noise}
\addplot[dash dot, line width=0.5mm] table[x=morozov_gammas, y=noise_datanorm] {MorozovData.dat};
\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}